---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE }
  library(dplyr)
  library(ggplot2)
  d=read.csv("activity.csv", stringsAsFactors = FALSE)
  x=complete.cases(d)
  final = tbl_df(d[x,])
  final = final %>% mutate(date=as.Date(date))
```


## What is mean total number of steps taken per day?
```{r echo=TRUE}
  noOfStepsByDate = final %>% group_by(date) %>% summarise(steps=sum(steps)) %>% mutate(dt=as.Date(date))
  ggplot(noOfStepsByDate, aes(x=dt, y=steps))+geom_col()
```

```{r echo=TRUE}
  summary(noOfStepsByDate$steps)
```


## What is the average daily activity pattern?
```{r echo=TRUE}
ggplot(final, aes(x=interval, y=steps))+geom_line()+stat_summary(geom="line",lwd=2, col="red")
```

```{r echo=TRUE}
noOfStepsByDate = final %>% group_by(interval) %>% summarise(steps=mean(steps))  %>% arrange(desc(steps))
noOfStepsByDate[1,]
```

## Imputing missing values
Missing Values Count
```{r echo=TRUE}
NumberOfMissingValues=count(tbl_df(d[!x,]))
as.numeric(NumberOfMissingValues)
```
```{r echo = TRUE}
avgStepsPerInterval = final %>% group_by(interval) %>% summarise(steps=mean(steps))
filledData=left_join(d[!x,],avgStepsPerInterval, by=c('interval')) %>% select(date,interval, steps.y)
colnames(filledData) <- c('date', 'interval', 'steps')
filledData <- tbl_df(filledData)
filledData= filledData %>% mutate(steps=as.integer(steps)) %>% mutate(date=as.Date(date))
all_data=full_join(filledData, final)
noOfStepsByDate1 = all_data %>% group_by(date) %>% summarise(steps=sum(steps)) %>% mutate(dt=as.Date(date))
  ggplot(noOfStepsByDate1, aes(x=dt, y=steps))+geom_col()

  summary(noOfStepsByDate$steps)
```
Comparing with the mean and median of the partial data, there is no significant difference after imputing missing data.
  
  
## Are there differences in activity patterns between weekdays and weekends?
```{r echo = TRUE}
  f_day = function(...){
    val = weekdays(...)
    #print(val)
    if (val == 'Sunday' || val == 'Saturday'){
      "Weekends"
    }
    else{
      "weekdays"
    }
  }
  all_data_group_by_week = all_data %>% mutate(weekday= ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "weekend", "weekday")) %>% group_by(weekday) %>% summarize(steps=sum(steps))
  
  
  
```

```{r echo=TRUE}
all_data1 = all_data %>% mutate(weekday= ifelse(weekdays(date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))


ggplot(all_data1, aes(x=interval, y=steps))+geom_line()+facet_grid(weekday ~ .)+stat_summary(geom="line",lwd=2, col="red")

```